Using Machine Learning and Constraint Programming to Solve the Relational-To-Ontology Mapping Problem

### 2017-08-11 stp/paper_aaai/main.tex
Checked. Ready for natalia to read and complete the missing parts

In our case, we would like this global model to account
not only for structural properties of the original data sources, but also to 
include the semantics which is usually implicitly present in the sources. In 
other words, we want to build a semantic model which describes the data sources 
in terms of concepts and relationships defined by an ontology. We can view this
semantic model as a graph with ontology classes as the nodes and ontology
properties as the links between the nodes \cite{Taheriyan2013}. Henceforth, we 
focus on a specific data integration problem: automatically mapping a new 
relational data source onto a user provided ontology.

@@

Firstly, we build an \emph{integration graph} which includes the 
attributes from the new source as well as ontology classes and properties. 
Secondly, we apply machine learning techniques to 
assign costs to its edges. 
Lastly, we use Constraint Programming to 
find a minimum cost Steiner Tree in the graph. The goal is to assign the costs 
on the edges in such a way that the resulting Steiner Tree is a valid semantic 
model for the new source.

@@

Section bla  presents bla bla... \ddg{TODO}

\section{Previous Work}
\ddg{rewrite, add more}
One of the standard approaches to solve the relational-to-ontology (\relonto{})
mapping problem is to apply Knowledge Representation and Reasoning (KR)
techniques. The main idea of this approach is to exploit constraints specified 
within relational schemata and ontologies. Several mapping generation
tools based on this approach already exist, for example, BootOX, MIRROR,
COMA++ \cite{pinkel2016rodi}. The issue is though that constraints may be 
inconsistent or absent completely \ddg{Why?}. We suggest to use ML techniques 
to overcome this hurdle.

@@

\section{Motivating Example}
\ddg{TODO}

\section{Problem Statement}
We follow the definitions proposed by Doal \etal{} \cite{doan2012principles}.

@@

An ontology $\mathcal{O}$ is a specification of classes of objects and their 
data properties and relationships between them. An example of an ontology is 
given 
in Fig. \ref{FIG:onto}. A Class in an ontology can have \emph{data properties} 
or 
\emph{class properties}. For example an ``Organisation'' has a ``Person'' 
(which is a class with its own properties), and it has a ``name'' (which 
is a data property). Also, a Class can be a subclass of another Class. For 
example ``Cities'', ``Countries'' or ``States'' are all ``Places''. A subclass 
inherits all the properties of the parent class.

@@

Our problem can be stated now the following way. We have an ontology 
$\mathcal{O}$ and a set of source descriptions $S_T = \{(s_1, m_1, \phi_1),..., 
(s_l, m_l, \phi_l)\}$.
Given a new source $s^\star$, we want to build a semantic model $m^\star$ and 
an attribute mapping function $\phi^\star$ such that 
$\delta^\star = (s^\star,m^\star,\phi^\star)$ is an \emph{appropriate} source 
description. We
use the term appropriate since there might be many such triples which are
well-formed source descriptions, but only one or a few will capture the intended
meaning of the source. Our goal is to automatically build $\delta^\star$ such 
that it minimizes the \emph{graph edit distance} (GED) from the semantic model 
$m^\star$ to the semantic model $m^\dag$ that the user considers correct.

@@

The alignment graph is a directed weighted graph $\mathcal{G_O} = 
(\mathcal{V_O},\mathcal{E_O})$ built on top 
of the known semantic models and expanded using the semantic types 
$\mathcal{L_O}$ and the domain ontology $\mathcal{O}$. Similar to a semantic 
model, $\mathcal{G_O}$ contains both class and data nodes. The links correspond 
to properties in  $\mathcal{O}$ and are weighted \cite{taheriyan2016learning}.

@@

The graph is weighted by a function $w_\mathcal{O} : \mathcal{E_O} \mapsto 
\mathbb{R}$ in 
such a way that links which are present in
the known semantic models have lower weights than those 
which are inferred from the ontology. As we will see int he next section, this 
makes links present in semantic models more attractive. \ddg{How exactly are 
they weighted?}

@@

\section{Steiner Tree Formulation}

@@

The solution $T^\star$ must satisfy the following constraints:
\begin{enumerate}
	\item $T^\star$ must be a subgraph of $\mathcal{I}_\mathcal{O}^{s^\star}$
	\item $T^\star$ must be a tree (connected acyclic graph)
	\item $\forall a \in \mathcal{A}_{s^\star}, a\in V^\star$
	\item $\forall a \in \mathcal{A}_{s^\star}, degree(a) = 1$
	\item $\forall n \in \mathcal{D_{G_O}}, degree(a) \in 
	\{0,2\}$
	\item $\forall n \in \mathcal{D_{G_O}} \cap V^\star, 
	degree(a) = 2$
\end{enumerate}

@@

To do this, we use the support of each of the obtained patters as a prize. If 
the tree contains a pattern, then its weight is automatically reduced by a the 
value of the support of that pattern. We will see in the next section how this 
information is integrated in the model.

\section{Modeling as a Constraint Optimization Problem}

@@

\subsection{Constraint Programming}
Constraint Programming (CP) allows the user to model a CSP or COP and give it 
to a 
solver that will find a solution to the CSP, or will report unsatisfiability if 
no such valuation exists. In the case of COPs, the solver will try to optimize 
the objective function and prove optimality as well.

@@

Since only using a backtracking algorithm would have a prohibitive cost, CP 
uses a combination of search and \emph{propagation}. The latter is achieved by 
\emph{propagating} the last decision made via specialized algorithms called 
\emph{propagators}. A propagator is a function $p : (v, D_v) \mapsto D_v'$ that 
takes a subset of the variables of the problem $v \subseteq \boldsymbol{v}$ and 
the domain of those variables $D_v$ and returns a new domain for thus variables 
such that $D_v' \sqsubseteq D$. That is, a propagator removes values form the 
domains of variables (when possible). For a propagator to be correct, it must 
not remove a value of the domain of a variable if such value could participate 
in a solution given the decisions already made.

Global constraints are higher order constraints that enforce a complicated 
constraint over a set ov variables. Typical examples of this is enforcing that 
a set of variables take different values, or that they describe a path a in a 
graph. Although it is possible to express these constraints with a conjunction 
of simple constraints, it is often the case that specific algorithms for these 
constraints perform substantially better than a decomposition approach (e.g. 
\cite{regin1994filtering}). A wide 
list of these global constraints can be found in \cite{beldiceanu2012global}.

\subsubsection{Example} Consider the toy problem $P$ over variables $a$ and $b$ 
such that $a \in \{2,3,4\}$, $b \in \{1,2,3\}$. Let $C = \{a + 2b \leq 4, a 
\neq b\}$ the constraints to be satisfied. If the search step decides that $a = 
2$, then (i) $a \neq b \Rightarrow b \in \{1,3\}$ and (ii) $a + 2b \leq 4  
\Rightarrow b \in \{1\}$. Thus we automatically know that $b = 1$ without never 
trying any other of its original possible values.

@@

Because there are global constraints in CP specialized in matching 
\cite{regin1994filtering}, we split the problem into two parts: (i) the 
\verb|steiner_tree| global constraint will only deal with the part of the 
integration graph that corresponds to the alignment graph, and the 
\verb|alldifferent| global constraint will deal with he matching part of the 
problem.

We use the following variables to represent the tree $T^\star$: Boolean 
variables $c_n,\forall n \in 
\mathcal{V_O}$, $c_e, \forall e \in	\mathcal{E_O}$ and an array of variables 
$match$ indexed by the set of attributes of $s^\star$. The combination of these 
sets fo variables define the value of $T^\star$: $c_n = \mathit{true}$ means 
that $n \in V^\star$, $c_e = \mathit{true}$ means that $e \in E^\star$ and 
$match[a] = d$ (for $a \in \mathcal{A}_{s^\star}$ and $d \in 
\mathcal{D_{G_O}}$) means that the edge $(a,d) \in 
\mathcal{M}_\mathcal{O}^{s^\star}$ is part of $E^\star$.

Additionally, for a given set of patterns $\mathcal{P}$ with a support function 
$w_\mathcal{P} : \mathcal{P} \mapsto \mathbb{R}$, we have a set of variables 
$c_p, \forall p \in \mathcal{P}$ that tells us whether a pattern $p$ appears in 
$T^\star$ or not.

\begin{flalign}
	& \text{Minimize~~} w_{STP} + w_{ADIFF} - w_{PAT}
	%\sum_{e \in \mathcal{E}_\mathcal{O}}c_e * 
	%w_\mathcal{I}(e) + \sum_{e \in \mathcal{M}_\mathcal{O}^{s^\star}} c_e * 
    %		w_\mathcal{I}(e)
	\label{EQ:obj}&&\\
	&\text{such that~~}  \nonumber&& \\
	& \textit{steiner\_tree}(\{c_n| n \in \mathcal{V_O}\},\{c_e| e \in 
	\mathcal{E_O}\},\mathcal{G_O},w_\mathcal{O},w_\mathit{STP})  \label{EQ:stp} 
	&&\\
	&\forall d \in \mathcal{D_{G_O}}, degree(d) \leq 1 \label{EQ:deg1}&&\\
	&\forall d \in \mathcal{D_{G_O}}, c_d \Leftrightarrow degree(d) = 1 
	\label{EQ:deg2}&&\\
	&\forall a \in \mathcal{A}_{s^\star}, match[a] \in \{ d | (a,d)\in 
	\mathcal{M}_\mathcal{O}^{s^\star} \} \label{EQ:matchdom}&&\\
	& \textit{alldifferent}(match) \label{EQ:alld}&& \\
	& \forall a \in \mathcal{A}_{s^\star}, c_{match[a]} = \mathit{true} 
	\label{EQ:map}&&\\
	& w_{\mathit{ADIFF}} = \sum_{(a,d) \in \mathcal{M}_\mathcal{O}^{s^\star}} 
	w_\mathcal{I}(~(a,d)~) * \llbracket match[a] = d\rrbracket 
	\label{EQ:matchcost}  &&\\
	& \forall p \in \mathcal{P}, \big(\forall e\in edges(p), c_e = 
	\mathit{true} \big) \Leftrightarrow c_p \label{EQ:patt}&&\\
	& w_{PAT} = \sum_{p\in\mathcal{P}} \mathit{support}[p] * c_p 
	\label{EQ:pcost} 
\end{flalign}

Eq. \ref{EQ:obj} is the objective function, we are trying to minimize the cost 
of $T^\star$ while collecting prizes for each pattern we use. Eq. \ref{EQ:stp} 
enforces that the solution 
$T^\star$ is indeed a tree defined by the $c_n$ and $c_e$ variables, subgraph 
of the alignment graph 
and of total weight $w_{STP}$. Equations \ref{EQ:deg1} and \ref{EQ:deg2} ensure 
that if a data node of the alignment graph is selected, then at most one edge 
reaches it (from the side of the alignment graph) and otherwise it is 
disconnected. Eq. \ref{EQ:matchdom} ensures that the domain of each variable in 
the array $match$ corresponds to a subset of data nodes of the alignment graph 
for which there is an edge connecting to the attribute at hand. Eq. 
\ref{EQ:alld} ensure that each attribute is mapped to exactly one data node of 
the alignment graph. Equations \ref{EQ:matchdom} and \ref{EQ:alld} ensure that 
the matching part of the problem is a bijection. Eq. \ref{EQ:map} ensures that 
if a data node of the alignment graph has been mapped to some attribute, then 
that data node must be in the solution tree, and vice-versa. Eq. 
\ref{EQ:matchcost} computes the cost $w_{ADIFF}$ of the matches.
Eq. \ref{EQ:patt} indicated that a pattern is used if and only if all its edges 
are selected in the tree. Eq. \ref{EQ:pcost} computes the prizes collected by 
using patterns.

\section{Results}

>>>>

In our case, we would like this global model to account
not only for structural properties of the original data sources, but also to 
include the semantics, which is usually implicitly present in the sources. In 
other words, we want to build a semantic model which describes the data sources 
in terms of concepts and relationships defined by an ontology 
\cite{Taheriyan2013}. Henceforth, we 
focus on a specific data integration problem: automatically mapping a new 
relational data source onto a user provided ontology.

\subsubsection{Example} Consider the situation where we have some simple 
relational database tables with columns $\langle$\textit{Surname, 
Workplace, Date, Location}$\rangle$, 
$\langle$\textit{Company, Event, Begin, Until, Location}$\rangle$ and 
$\langle$\textit{Name, D.O.B, Location}$\rangle$. We do not \textit{a priori} 
know whether ``Date'' is the date of birth or the date a persons started 
working in the first table; or whether the ``Location'' on the second table 
refers to the location of the event, or the address of the company. 
If we are given an ontology like the one in Fig. \ref{FIG:onto}, we would like 
to automatically map the new data sources to classes of the 
ontology.$\square$

@@

Firstly, we build an \emph{integration graph} which includes the 
attributes from the new source as well as ontology classes and properties. 
Secondly, we apply machine learning techniques to 
assign costs to its edges. 
Lastly, we use Constraint Programming (CP) to 
find a minimum cost Steiner Tree in the graph. The goal is to assign the costs 
on the edges in such a way that the resulting Steiner Tree is a valid semantic 
model for the new source.

@@

Section \ref{SEC:pw} presents the previous work done to achieve this task. 
Section \ref{SEC:problem} formally states the problem. 
In Section \ref{SEC:ML} we present how we convert the information given into a 
useful data representation. 
Section \ref{SEC:STP} shows how we model the problem as a STP, 
whereas Section \ref{SEC:CP} discusses how we implement the model in CP. 
Section \ref{SEC:Res} shows our results.

\section{Previous Work \label{SEC:pw}}
\ddg{rewrite, add more}
One of the standard approaches to solve the relational-to-ontology (\relonto{})
mapping problem is to apply Knowledge Representation and Reasoning (KR)
techniques. The main idea of this approach is to exploit constraints specified 
within relational schemata and ontologies. Several mapping generation
tools based on this approach already exist, for example, BootOX, MIRROR,
COMA++ \cite{pinkel2016rodi}. The issue is though that constraints may be 
inconsistent or absent completely \ddg{Why?}. We suggest to use ML techniques 
to overcome this hurdle.

@@

\section{Problem Statement \label{SEC:problem}}
We follow the definitions proposed by Doal \etal{} \cite{doan2012principles}.

@@

An ontology $\mathcal{O}$ is a specification of classes of objects, their 
data properties and relationships between them. An example of an ontology is 
given 
in Fig. \ref{FIG:onto}. A Class in an ontology can have \emph{data properties} 
or 
\emph{class properties}. For example an ``Organisation'' has a ``Person'' 
(which is a class with its own properties), and it has a ``name'' (which 
is a data property). Also, a Class can be a subclass of another Class. For 
example ``Cities'' and ``States'' are both ``Places''. A subclass 
inherits all the properties of the parent class.

@@

Our problem can be stated now the following way. We have an ontology 
$\mathcal{O}$ and a set of source descriptions $S_T = \{(s_1, m_1, \phi_1),..., 
(s_l, m_l, \phi_l)\}$.
Given a new source $s^\star$, we want to build a semantic model $m^\star$ and 
an attribute mapping function $\phi^\star$ such that 
$\delta^\star = (s^\star,m^\star,\phi^\star)$ is an \emph{appropriate} source 
description. We
use the term ``appropriate'' since there might be many such triples which are
well-formed source descriptions, but only one or a few will capture the intended
meaning of the source. Our goal is to automatically build $\delta^\star$ such 
that it minimizes the \emph{graph edit distance} (GED) from the semantic model 
$m^\star$ to the semantic model $m^\dag$ that the user considers correct.

@@

The alignment graph is a directed weighted graph $\mathcal{G_O} = 
(\mathcal{V_O},\mathcal{E_O})$ built on top 
of the known semantic models and expanded using the semantic types 
$\mathcal{L_O}$ and the ontology $\mathcal{O}$. Similar to a semantic 
model, $\mathcal{G_O}$ contains both class and data nodes. The links correspond 
to properties in  $\mathcal{O}$ and are weighted \cite{taheriyan2016learning}.

@@

The graph is weighted by a function $w_\mathcal{O} : \mathcal{E_O} \mapsto 
\mathbb{R}$ in 
such a way that edges which are present in
the known semantic models have lower weights than those 
which are inferred from the ontology. As we will see int he next section, this 
makes edges present in semantic models more attractive. \ddg{How exactly are 
they weighted?}

@@

\section{Steiner Tree Formulation \label{SEC:STP}}

@@

The solution $T^\star$ must satisfy the following constraints:
\begin{enumerate}
	\item $T^\star$ must be a subgraph of $\mathcal{I}_\mathcal{O}^{s^\star}$
	\item $T^\star$ must be a tree (connected acyclic graph)
	\item $\forall a \in \mathcal{A}_{s^\star}, a\in V^\star$
	\item $\forall a \in \mathcal{A}_{s^\star}, degree(a) = 1$
	%\item $\forall n \in \mathcal{D_{G_O}}, degree(a) \in 
	%\{0,2\}$
	\item $\forall n \in \mathcal{D_{G_O}} \cap V^\star, 
	degree(a) = 2$
\end{enumerate}

@@

To do this, we use the support of each of the obtained patterns as a prize. If 
the tree contains a pattern, then its weight is automatically reduced by a the 
value of the support of that pattern. We will see in the next section how this 
information is integrated in the model.

\section{Modeling as a Constraint Optimization Problem \label{SEC:CP}}

@@

\subsection{Constraint Programming}
Constraint Programming (CP) allows the user to model a CSP or COP and give it 
to a 
solver that will find a solution to the CSP, or will report unsatisfiability if 
no solution exists. In the case of COPs, the solver will try to optimize 
the objective function and prove optimality as well.

@@

Since only using a backtracking algorithm would have a prohibitive cost, CP 
uses a combination of search and \emph{propagation}. The latter is achieved by 
\emph{propagating} the last decision made via specialized algorithms called 
\emph{propagators}. A propagator is a function $p : (v, D_v) \mapsto D_v'$ that 
takes a subset of the variables of the problem $v \subseteq \boldsymbol{v}$ and 
the domain of those variables $D_v$ and returns a new domain for these 
variables 
such that $D_v' \sqsubseteq D$. That is, a propagator removes values form the 
domains of variables (when possible). For a propagator to be correct, it must 
not remove a value of the domain of a variable if such value could participate 
in a solution given the decisions already made.

\emph{Global constraints} are higher order constraints that enforce a 
complicated 
constraint over a set of variables. Typical examples of this is enforcing that 
a set of variables take different values, or that they describe a path a in a 
graph. Although it is possible to express these constraints with a conjunction 
of simple constraints, it is often the case that specific algorithms for these 
constraints perform substantially better than a decomposition approach (e.g. 
\cite{regin1994filtering}). A wide 
list of these global constraints can be found in \cite{beldiceanu2012global}.

\subsubsection{Example} Consider the toy problem $P$ over variables $a$ and $b$ 
such that $a \in \{2,3,4\}$, $b \in \{1,2,3\}$. Let $C = \{a + 2b \leq 4, a 
\neq b\}$ be the constraints. If the search step decides that $a = 
2$, then $(i)$ $a \neq b \Rightarrow b \in \{1,3\}$ and $(ii)$ $a + 2b \leq 4  
\Rightarrow b \in \{1\}$. Thus we automatically know that $b = 1$ without never 
trying any other of its original possible values. $\square$

@@

Because there are global constraints in CP specialized in matching 
\cite{regin1994filtering}, we split the problem into two parts: the 
\verb|steiner_tree| global constraint will only deal with the part of the 
integration graph that corresponds to the alignment graph, and the 
\verb|alldifferent| global constraint will deal with he matching part of the 
problem.

We use the following variables to represent the tree $T^\star$: Boolean 
variables $c_n,\forall n \in 
\mathcal{V_O}$, $c_e, \forall e \in	\mathcal{E_O}$ and an array of variables 
$match$ indexed by the set of attributes of $s^\star$. The combination of these 
sets fo variables define the value of $T^\star$: $c_n = \mathit{true}$ means 
that $n \in V^\star$, $c_e = \mathit{true}$ means that $e \in E^\star$ and 
$match[a] = d$ (for $a \in \mathcal{A}_{s^\star}$ and $d \in 
\mathcal{D_{G_O}}$) means that the edge $(a,d) \in 
\mathcal{M}_\mathcal{O}^{s^\star}$ is part of $T^\star$.

Additionally, for a given set of patterns $\mathcal{P}$ with a support function 
$w_\mathcal{P} : \mathcal{P} \mapsto \mathbb{R}$, we have a set of Boolean 
variables 
$c_p, \forall p \in \mathcal{P}$ that tells us whether a pattern $p$ appears in 
$T^\star$ or not.
The model is presented below.

\begin{flalign}
	& \text{Minimize~~} w_{STP} + w_{ADIFF} - w_{PAT}
	%\sum_{e \in \mathcal{E}_\mathcal{O}}c_e * 
	%w_\mathcal{I}(e) + \sum_{e \in \mathcal{M}_\mathcal{O}^{s^\star}} c_e * 
    %		w_\mathcal{I}(e)
	\label{EQ:obj}&&\\
	&\text{such that~~}  \nonumber&& \\
	& \textit{steiner}(\{c_n| n \in \mathcal{V_O}\},\{c_e| e \in 
	\mathcal{E_O}\},\mathcal{G_O},w_\mathcal{O},w_\mathit{STP})  \label{EQ:stp} 
	&&\\
	&\forall d \in \mathcal{D_{G_O}}, degree(d) \leq 1 \label{EQ:deg1}&&\\
	&\forall d \in \mathcal{D_{G_O}}, c_d \Leftrightarrow degree(d) = 1 
	\label{EQ:deg2}&&\\
	&\forall a \in \mathcal{A}_{s^\star}, match[a] \in \{ d | (a,d)\in 
	\mathcal{M}_\mathcal{O}^{s^\star} \} \label{EQ:matchdom}&&\\
	& \textit{alldifferent}(match) \label{EQ:alld}&& \\
	& \forall a \in \mathcal{A}_{s^\star}, c_{match[a]} = \mathit{true} 
	\label{EQ:map}&&\\
	& w_{\mathit{ADIFF}} = \sum_{(a,d) \in \mathcal{M}_\mathcal{O}^{s^\star}} 
	w_\mathcal{I}(~(a,d)~) * \llbracket match[a] = d\rrbracket 
	\label{EQ:matchcost}  &&\\
	& \forall p \in \mathcal{P}, \big(\forall e\in edges(p), c_e = 
	\mathit{true} \big) \Leftrightarrow c_p \label{EQ:patt}&&\\
	& w_{PAT} = \sum_{p\in\mathcal{P}} w_\mathcal{P}(p)* c_p 
	\label{EQ:pcost} 
\end{flalign}

Eq. \ref{EQ:obj} is the objective function: we are minimizing the cost 
of $T^\star$ while collecting prizes for each pattern we use. Eq. \ref{EQ:stp} 
enforces that the solution 
$T^\star$ is indeed a tree defined by the $c_n$ and $c_e$ variables, subgraph 
of the alignment graph 
and of total weight $w_{STP}$. Equations \ref{EQ:deg1} and \ref{EQ:deg2} ensure 
that if a data node of the alignment graph is selected, then at most one edge 
reaches it (from the side of the alignment graph) and otherwise it is 
disconnected. Eq. \ref{EQ:matchdom} ensures that the domain of each variable in 
the array $match$ corresponds to a subset of data nodes of the alignment graph 
for which there is an edge connecting to the attribute at hand. Eq. 
\ref{EQ:alld} ensure that each attribute is mapped to exactly one data node of 
the alignment graph. Equations \ref{EQ:matchdom} and \ref{EQ:alld} ensure that 
there is exactly one data node matched to an attribute, which has to be one of 
the ones for which an edge existed. Eq. \ref{EQ:map} ensures that 
if a data node of the alignment graph has been mapped to some attribute, then 
that data node must be in the solution tree, and vice-versa. Eq. 
\ref{EQ:matchcost} computes the cost $w_{ADIFF}$ of the matches.
Eq. \ref{EQ:patt} indicated that a pattern is used if and only if all its edges 
are selected in the tree. Eq. \ref{EQ:pcost} computes the prizes collected by 
using patterns.

\section{Results \label{SEC:Res}}
